---
title: "README"
output: html_document
---

    
# Miscellaneous notes

     I fixed the issue with regards to the source function, but I remember
     you said I should rather make it a function? 
    
    In the preparation file, when it wants the data to be read in table 
    format, it says <'file' must be a character string or connection >
    
    Do you know how I can fix it?
    
    With regards to the U.S. data, everything works fine up until line 56
    where I try and merge the two series together.
    
    I've now tried with Euro data as well, but I get the familiar response
    that I get with the other prepare files as well:
    
    awm.data <- read.table(awm.file,header=TRUE)[c('DATE','YER','STN','HEX','HICP'),]
    Error in read.table(awm.file, header = TRUE) : 
         'file' must be a character string or connection
        
        THis seems to happen across all prepare files.
    
  
              


# Sourcing your functions:

```{r}
library(tidyverse)
listSources <- list.files("Code", full.names = TRUE)
listSources %>% as.list() %>% map(~source(.))
```



# Aim

THe aim of this paper will be to estimate a natural rate of interest for South Africa. 

Changes

To add
DM yields lower, em yields higher 
Elevated yield differential and then well see an outflow of the rand
Implications for carry trade 
Want to see if rates are too high and too low
Check deutshe bank carry index-- investable index

# Techniques used

  *Univariate time series estimates
  ---------------------------------
  Multi-year averages
  Ten year moving averages of ex post real st i-rates
  
  Time series filtering techniques
  --------------------------------
  Hodrik-Prescott filter
  Bandpass filter
  Unobserved components stochastic volatility model
  
  *Multivariate Kalman filter estimates
  -------------------------------------
  LAUBACH WILLIAMS MODEL

# Data

  For replication results
  -----------------------
  
  Schularik and Taylor dataset
  For each economy:
          - real GDP
          - inflation
              -annualized quarterly growth rate of the specified consumer price series
          - short-term nominal interest rate
              -expressed on a 365 day annualized basis
          - procedure to compute inflation expectations
              - in order to calculate the ex ante real short-term interest rate
    
  US DATA
  -------
  
  - Real GDP and core PCE
  - ST i-rate is the annualized nominal federal funds rate
  - All US data from the St Louis FRED
  
  CANADA DATA - all data from Statistics Canada except real GDP
  -----------
  
  - Real GDP from IMF's IFS
  - ST i-rate ---> Bank of Canada's overnight target rate
              ---> taken as the end of period value for each month
              ---> aggregated to quarterly frequency
              ---> prior to 2001, LW used BoC's bank rate as the i-rate
  - Inlfation ---> BoC's consumer price index
              ---> prior to 1984: use CPI containing all items
  
  EU DATA - all data from EU's Area-Wide Model
  ------- - available from the Euro Area Business Cycle Network
  
  - CPI beginning in 1988; total price index prior to 1988
  - nominal ST i-rate is the 3 month rate
  
  UK DATA
  -------
  
  - GDP data - office of national statistics (ONS)
  - Inflation - core CPI
              - prior to 1970 used all items CPI; both from OECD
  - ST i-rate - Bank of ENgland's official bank rate
    
    
Remember to delete all data once the project is finished.
    
# Packages used

    tis
    mFilter
    nloptr
    HelpersMG
    seasonalview
    lubridate
    
# PAPER

## Introduction

   *Outline what the skeleton of the paper is going to be.
   Discuss the rationale behind estimating a natural rate of interest
    - Why do we need one?
    - Policymaking- must they acknowledge the natural rate or try implement policies
                    that are not heavily affected by the natural rate
    - Discuss Laubach-Williams model
    - Add a very brief, concise discussion on the theoretical and then empirical processes
      to estimate the natural rate of interest
   
## Methodology

  # What needs to be replicated:
     -natural rate of interest
     -potential GDP
     -potential GDP trend growth
     - for following countries: United States, Canada, Euro area, UK
     
  # And then repeat the same process for South Africa
     -maybe for other Southern African states if there is the data/if its necessary

list code

ctrl alt i

```{r Initialize}
L <- list.files("Code", full.names = TRUE)
for(i in 1:length(L)){
  source(L(i))
}

library(rmsfuns)
load.pkg(c("PerformanceAnalytics", "tidyverse", ))
```


```{r Data Import}

getFRED()

```

```{r utilities}

#------------------------------------------------------------------------------#
# File:        utilities.R
#
# Description: This file contains basic functions that will be used throughout
#              HLW.
#------------------------------------------------------------------------------#

shiftQuarter <- function(original.start,shift){
#################################################################
# This function takes in a (year,quarter) date in time series format
# and a shift number, and returns the (year,quarter) date corresponding
# to the shift. Positive values of shift produce leads and negative values
# of shift produce lags.
# For example, entering 2014q1 with a shift of -1 would return 2013q4.
# Entering 2014q1 with a shift of 1 would return 2014q2.
# In each case, the first argument of the function must be entered as
# a two-element vector, where the first element corresponds to the year
# and the second element corresponds to the quarter.
# For example, Q12014 must be entered as "c(2014,1)".
################################################################    

# Leads (positive values of shift)
    if (shift > 0) {
        new.start = c(0,0)
        sum = original.start[2] + shift
    
        # Get the year value
        if (sum <= 4) {
            new.start[1] = original.start[1]
        }
        else {
            new.start[1] = original.start[1] + ceiling(sum/4) - 1
        }

        # Get the quarter value
        if (sum %% 4 > 0) {
            new.start[2] = sum %% 4
        }
        else {
            new.start[2] = sum %% 4 + 4
        }
    }

# Lags (negative values of shift)
    else {
        new.start = c(0,0)
        diff = original.start[2] - abs(shift)
    
        # Get the year value
        if (diff > 0) {
            new.start[1] = original.start[1]
        }
        else {
            new.start[1] = original.start[1] - (1 + floor(abs(diff)/4))
        }

        # Get the quarter value
        if (diff %% 4 > 0) {
            new.start[2] = diff %% 4
        }
        else {
            new.start[2] = diff %% 4 + 4
        }
    }
        
return(new.start)}


shiftMonth <- function(original.start,shift){
#################################################################
# This function takes in a (year,month) date in time series format
# and a shift number, and returns the (year,month) date corresponding
# to the shift. Positive values of shift produce leads and negative values
# of shift produce lags.
# For example, entering 2014m1 with a shift of -1 would return 2013m12.
# Entering 2014m1 with a shift of 1 would return 2014m2.
# In each case, the first argument of the function must be entered as
# a two-element vector, where the first element corresponds to the year
# and the second element corresponds to the month.
# This function is analogous to shiftQuarter().
################################################################    

# Leads (positive values of shift)
    if (shift > 0) {
        new.start = c(0,0)
        sum = original.start[2] + shift
    
        # Get the year value
        if (sum <= 12) {
            new.start[1] = original.start[1]
        }
        else {
            new.start[1] = original.start[1] + ceiling(sum/12) - 1
        }

        # Get the month value
        if (sum %% 12 > 0) {
            new.start[2] = sum %% 12
        }
        else {
            new.start[2] = sum %% 12 + 12
        }
    }

# Lags (negative values of shift)
    else {
        new.start = c(0,0)
        diff = original.start[2] - abs(shift)
    
        # Get the year value
        if (diff > 0) {
            new.start[1] = original.start[1]
        }
        else {
            new.start[1] = original.start[1] - (1 + floor(abs(diff)/12))
        }

        # Get the month value
        if (diff %% 12 > 0) {
            new.start[2] = diff %% 12
        }
        else {
            new.start[2] = diff %% 12 + 12
        }
    }
        
return(new.start)}


getFRED <- function(url, freq = "Quarterly") {
##########################################################################################
# This function downloads data from FRED. It returns quarterly data.
# User must provide the FRED url.
########################################################################################### 
    # Download the data from FRED
    
    #download.file(url, destfile = 'FREDtemp.txt', method = "wget")
    #FREDraw <- readLines('FREDtemp.txt')
    
    txt.file.name <- paste0("rawData/",substr(url, regexpr('[a-zA-z0-9]*.txt',url),1000))
    if (!file.exists(txt.file.name)){
        # Download the data from FRED
        #download.file(url, destfile = 'FREDtemp.txt', method = "wget")
        system(paste0('wget --no-check-certificate "', url, '"'))
        system(paste('mv',substr(url, regexpr('[a-zA-z0-9]*.txt',url),1000),txt.file.name))
    }
    FREDraw <- readLines(txt.file.name) 

    # Frequency
    freq.FRED <- gsub(' ', '',substr(FREDraw[which(regexpr('Frequency', FREDraw)==1)],
                                     (nchar('Frequency')+2),100))    

    # Where does the data start
    datastart = which(gsub(' ', '',FREDraw)=='DATEVALUE') - 2

    #data <- read.table('FREDtemp.txt', skip = datastart, header = TRUE)
    data <- read.table(txt.file.name, skip = datastart, header = TRUE)

    first.year  <- as.numeric(format(as.Date(data$DATE[1]),'%Y'))
    first.month <- as.numeric(format(as.Date(data$DATE[1]),'%m'))
    
    # Adjust frequency
    if (freq.FRED == 'Quarterly'){
        first.q  <- (first.month-1)/3 + 1
        data.tis <- tis(data$VALUE, start = c(first.year, first.q), tif = 'quarterly')
    } else if (freq.FRED == 'Monthly') {
        data.tis <- tis(data$VALUE, start = c(first.year, first.month), tif = 'monthly')
    }

    # Convert frequency
    if (freq.FRED == 'Monthly' & freq == 'Quarterly') {
        data.tis <- convert(data.tis, tif = 'quarterly', method = 'constant', observed. = 'averaged')
    }

    return(data.tis)
} 


splice <- function(s1, s2, splice.date, freq) {
##########################################################################################
# This function splices two series, with the series s2 beginning at splice.date
# and extended back using the growth rate at the splice.date times series s1
# The freq argument accepts two values - 'quarterly' and 'monthly' -
# but it could be modified to take more.
##########################################################################################    
    t <- splice.date #renaming for convenience
    if (freq == "quarterly" | freq == "Quarterly") {
        t.minus.1 <- shiftQuarter(t,-1)
    }
    else if (freq == "monthly" | freq == "Monthly") {
        t.minus.1 <- shiftMonth(t,-1)
    }
    else { stop("You must enter 'quarterly' or 'monthly' for freq.") }
    ratio <- as.numeric(window(s2,start = t, end = t)/
                        window(s1,start = t, end = t))

    return(mergeSeries(ratio*window(s1,end = t.minus.1),window(s2, start = t)))
}


gradient <- function(f, x, delta = x * 0 + 1.0e-5) {
##########################################################################################
# This function computes the gradient of a function f given a vector input x.
##########################################################################################   
    g <- x * 0
    for (i in 1:length(x)) {
        x1 <- x
        x1[i] <- x1[i] + delta[i]
        f1 <- f(x1)
        x2 <- x
        x2[i] <- x2[i] - delta[i]
        f2 <- f(x2)
        g[i] <- (f1 - f2) / delta[i] / 2
    }
    return(g)
}

```
 
 # This prepares the following functions:
   - shiftQuarter
   - shiftMonth
   - getFRED
   - splice
   - gradient
   
 
```{r}

```



## Lit Review

 The goal here will be to arrive at a complete definition on what the natural rate of interest is, one that
 is firmly anchored in the literature.
 Provide a proper disection of the Laubach.
 
 Natural rate of interest==> intersection of IS curve and potential GDP line. 
       - To add ==> explain a shift in the IS/potential GDP and its effect on the natural rate 
       - What affects potential GDP?
       - What affects the real interest rate and thus the natural rate?
       - LT/ST perspectives??
        
NB question- why do we need to know what the natural rate of interest is?
        -> then explain policies 
 

## Results

## COnclusion