---
title: "README"
output: html_document
---

# Aim

THe aim of this paper will be to estimate a natural rate of interest for South Africa. 


# Techniques used

  *Univariate time series estimates
  ---------------------------------
  Multi-year averages
  Ten year moving averages of ex post real st i-rates
  
  Time series filtering techniques
  --------------------------------
  Hodrik-Prescott filter
  Bandpass filter
  Unobserved components stochastic volatility model
  
  *Multivariate Kalman filter estimates
  -------------------------------------
  LAUBACH WILLIAMS MODEL

# Data

  For replication results
  -----------------------
  
  Schularik and Taylor dataset
    
    
    
Remember to delete all data once the project is finished.
    
# Packages used

    ....
    
# Miscellaneous notes

    * Note when running this function, BigFunc, it takes roughly 20 minutes.
    
    * I have issues having my model converge
    
    * I am stuck with code in section xxx, please check
    

# PAPER

## Introduction

   *Outline what the skeleton of the paper is going to be.
   Discuss the rationale behind estimating a natural rate of interest
    - Why do we need one?
    - Policymaking- must they acknowledge the natural rate or try implement policies
                    that are not heavily affected by the natural rate
    - Discuss Laubach-Williams model
    - Add a very brief, concise discussion on the theoretical and then empirical processes
      to estimate the natural rate of interest
   
## Methodology

list code

ctrl alt i

```{r Initialize}
L <- list.files("Code", full.names = TRUE)
for(i in 1:length(L)){
  source(L(i))
}

library(rmsfuns)
load.pkg(c("PerformanceAnalytics", "tidyverse", ))
```


```{r Data Import}

Data <- read_rds("Data/ ")

```

```{r Returns create}

```




## Lit Review

 The goal here will be to arrive at a complete definition on what the natural rate of interest is, one that
 is firmly anchored in the literature.
 Provide a proper disection of the Laubach.
 
 Natural rate of interest==> intersection of IS curve and potential GDP line. 
       - To add ==> explain a shift in the IS/potential GDP and its effect on the natural rate 
       - What affects potential GDP?
       - What affects the real interest rate and thus the natural rate?
       - LT/ST perspectives??
        
NB question- why do we need to know what the natural rate of interest is?
        -> then explain policies 
 

## Results

## COnclusion